\documentclass[12pt]{article} 

\usepackage[%
  top    = 2.5cm,
  bottom = 2.5cm,
  left   = 2cm,
  right  = 2cm,
]{geometry}

\newcommand\customBlank{%
  \rule{2cm}{0.15mm} \ %
}

\title{Past Simple vs Present Perfect}
\author{Philippe Fanaro}
\date{March 26th, 2026}

\begin{document}

\maketitle

\section{Definitions}

We use the \textbf{past simple} to talk about actions that happened at a specific time in the past. The exact time is often mentioned or implied. For example:

\begin{itemize}
  \item I \textbf{visited} Paris last year.
  \item She \textbf{finished} her homework yesterday.
  \item They \textbf{moved} to a new house in 2010.
\end{itemize}

In contrast, we use the \textbf{present perfect} to talk about actions that happened at an unspecified time in the past or that have relevance to the present moment. The exact time is not important or not mentioned. However, we can still use adverbs for specifying the time more, if it is more relevant to the present (the time will still remain vague).

\begin{itemize}
  \item I \textbf{have visited} Paris (before).
  \item She \textbf{has finished} her homework (earlier).
  \item They \textbf{have moved} to a new house (already).
\end{itemize}

When we use the \textbf{present perfect}, we often want to emphasize the experience or the result of the action rather than the specific time when it occurred.

\begin{itemize}
  \item I \textbf{have lived} here for five years. (emphasizing the experience)
  \item I \textbf{have visited} Paris, and it was amazing! (emphasizing the experience)
  \item She \textbf{has finished} her homework, so now she can relax. (emphasizing the result)
  \item They \textbf{have moved} to a new house, and they love it! (emphasizing the experience and result)
\end{itemize}

\section{Summary}

\begin{itemize}
  \item Use the \textbf{past simple} for actions that happened at a specific time in the past.
  \item Use the \textbf{present perfect} when:
    \begin{itemize}
      \item The time is vague, nor mentioned, or not important.
      \item When even though, the time might be more speficied, the information is more relevant to the present.
    \end{itemize}
\end{itemize}

\section{Exercises}

Fill in the blanks with the correct form of the verb (past simple or present perfect):

\begin{enumerate}
  \item I \customBlank (to be) to New York before.
  \item I \customBlank (to be) in New York last year.
  \item I \customBlank (to eat) an apple this morning.
  \item I \customBlank (to eat) apples earlier.
  \item We \customBlank (to watch) 3 episodes of Hikaru no Go already.
  \item We \customBlank (to play) Baduk only once. Never again!
  \item He \customBlank (to be) in Seoul for 2 months. (Still there)
  \item He \customBlank (to be) in Seoul for 2 months. (Not there anymore)
  \item For how long \customBlank (to have) a dog? (The dog is still alive)
  \item For how long \customBlank (to have) a dog? (The dog is not alive anymore)
\end{enumerate}

\end{document}